### 2020-12-11 ch02_scientific_background/scientific_background.tex
Finished section on climate sensitivity

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/gregory_regression.pdf}
  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} global and
    annual mean near-surface air temperature $\Delta T$ for 150 years of a
    simulation with an abrupt quadrupling of the atmospheric \ac{CO2}
    concentration (4x\acs{CO2}) for the \acs{CMIP}5 multi-model mean (blue
    squares). To account for energy leakage and model drift, a linear fit of
    the corresponding pre-industrial control run is subtracted from the
    4x\acs{CO2} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope of
    the linear regression (black line) corresponds to the climate feedback
    parameter $\lambda$ and the y-intercept corresponds to the radiative
    forcing $F_\text{4x}$. These can be used to calculate the \ac{ECS} with
    the Gregory regression method following to \cref{eq:02:ecs}
    \autocite{Gregory2004}.}
  \label{fig:02:gregory_regression}
\end{figure}

For this reason, the Equilibrium Climate Sensitivity is commonly approximated
with the \acf{ECS}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(4x\acs{CO2}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a 4x\acs{CO2}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ \vs{} the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
4x\acs{CO2} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the 4x\acs{CO2} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the y-intercept of this linear regression
($F_\text{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \text{ECS} = - \frac{F_\text{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

Although commonly used in the literature, the \ac{ECS} is known to be only an
approximation of the true Equilibrium Climate Sensitivity. One major reason for
this is the state and time dependence of the global feedbacks
\autocite{Knutti2015, Knutti2017}. As
\cref{fig:02:gregory_regression_different_years} shows, the slope in the
Gregory regression is not constant, but rather decreases over time when a long
running 4x\acs{CO2} simulation with more than 1000 model years is considered.
As a result, the climate feedback parameter $\lambda$ decreases over time,
resulting in a higher \ac{ECS}. Major reasons for this are temperature
dependencies of the feedbacks, atmospheric and oceanic adjustments over time,
changing warming patterns over time, non-additive feedbacks and dependencies on
the type and magnitude of the external forcings \autocite{Knutti2017}. All in
all, this demonstrates the limits of the linear feedback framework introduced
in \cref{subsec:02:mathematical_framework_feedbacks} that is not capable of
describing non-linear effects. A second major reason for the discrepancies
between the Equilibrium Climate Sensitivity and \ac{ECS} is the use of a
4x\acs{CO2} instead of a 2x\acs{CO2} simulation. The factor of $2$ in the
denominator of \cref{eq:02:ecs} only partly compensates this difference since
the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
concentration \autocite{Huang2014}.

@@

A further metric describing the climate sensitivity is the \ac{TCR}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the climate system but describes the transient response of an evolving climate.
\ac{TCR} is defined as the change in the annual mean \ac{GSAT} at the time of
\ac{CO2} doubling in a simulation with a \pct{1} per year increase of the
atmospheric \ac{CO2} concentration (1\%\acs{CO2}) averaged over a 20 year
period centered at the time of the doubling (years \range{61}{80} when the
first year is indexed with 1) \autocite{Bindoff2013}. To account for model
drift, the annual mean changes in \ac{GSAT} are calculated relative to a
corresponding pre-industrial control simulation smoothed with a linear fit that
considers 140 model years (length of the 1\%\acs{CO2} simulation).

- ocean heat uptake (TCR)

- residual heat -> 90\% go into ocean (Knutti et al. 2017)

- equations for TCR

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/gregory_regression.pdf}
  \caption{Global and annual mean net \acf{TOA} radiation $N$ \vs{} the change
    in global and annual mean near-surface air temperature $\Delta T$ for 150
    years of a simulation with an abrupt quadrupling of the atmospheric
    \ac{CO2} concentration (\nxcotwo{4}) for the \acs{CMIP}5 multi-model mean
    (blue circles). To account for energy leakage and model drift, a linear
    fit of the corresponding pre-industrial control run is subtracted from the
    \nxcotwo{4} simulation. As given by \cref{eq:02:N_vs_delta_t}, the slope
    of the linear regression (black line) corresponds to the climate feedback
    parameter $\lambda$ and the y-intercept corresponds to the radiative
    forcing $F_\text{4x}$. These can be used to calculate the \acf{ECS} with
    the Gregory regression method according to \cref{eq:02:ecs}
    \autocite{Gregory2004}. \acs{ECS} is equivalently given by the x-intercept
    of the linear regression line divided by $2$. Here, $\text{ECS} = 3.14
    \unit{K}$.}
  \label{fig:02:gregory_regression}
\end{figure}

For this reason, the Equilibrium Climate Sensitivity is commonly approximated
with the \acf{ECS}, which can be derived from only 150 model years of a
simulation with an abrupt quadrupling of the atmospheric \ac{CO2} concentration
(\nxcotwo{4}) \autocite{Gregory2004}. The basis of the definition of \ac{ECS}
is the simple energy balance model introduced in
\cref{subsec:02:mathematical_framework_feedbacks}. Assuming radiative
equilibrium ($N = 0$), \cref{eq:02:N,eq:02:linear_response} imply
\begin{equation}
  \Delta T = -\frac{F}{\lambda}.
  \label{eq:02:delta_t_eq}
\end{equation}
Thus, the change in \ac{GSAT} in radiative equilibrium can be easily calculated
from the external forcing $F$ and the climate feedback parameter $\lambda$. The
steady state values for $F$ and $\lambda$ can be estimated from a \nxcotwo{4}
simulation that is not in equilibrium by extrapolation with a linear regression
\autocite{Gregory2004}. For this method, the so-called \emph{Gregory
  regression}, the global and annual mean net \ac{TOA} radiation $N$ \vs{} the
change in the annual mean \ac{GSAT} $\Delta T$ for all 150 years of the
\nxcotwo{4} run are plotted (see \cref{fig:02:gregory_regression}). To account
for energy leakage and remove any model drift present in the control climate, a
linear fit of the corresponding pre-industrial control run is subtracted from
the \nxcotwo{4} simulation beforehand \autocite{Andrews2012}. Since the
combination of \cref{eq:02:N,eq:02:linear_response} yields
\begin{equation}
  N = F + \lambda \cdot \Delta T,
  \label{eq:02:N_vs_delta_t}
\end{equation}
$F$ is now given by the y-intercept of this linear regression
($F_\text{4x}$) and $\lambda$ by its slope. Thus, the \ac{ECS} is given by
\begin{equation}
  \text{ECS} = -\frac{F_\text{4x}}{2 \lambda}.
  \label{eq:02:ecs}
\end{equation}
The factor of $2$ in the denominator accounts for the fact the traditional
Equilibrium Climate Sensitivity is defined for an abrupt \ac{CO2} doubling,
whereas here a simulation with an abrupt quadrupling is considered.

@@

Although commonly used in the literature, the \ac{ECS} is known to be only an
approximation of the true Equilibrium Climate Sensitivity. One major reason for
this is the state and time dependence of the global feedbacks
\autocite{Knutti2015, Knutti2017}. As
\cref{fig:02:gregory_regression_different_years} shows, the slope in the
Gregory regression is not constant, but rather decreases over time when a long
running \nxcotwo{4} simulation with more than 1000 model years is considered.
As a result, the climate feedback parameter $\lambda$ decreases over time,
resulting in a higher \ac{ECS}. Major reasons for this are temperature
dependencies of the feedbacks, atmospheric and oceanic adjustments over time,
changing warming patterns over time, non-additive feedbacks and dependencies on
the type and magnitude of the external forcings \autocite{Knutti2017}. All in
all, this demonstrates the limits of the linear feedback framework introduced
in \cref{subsec:02:mathematical_framework_feedbacks} that is not capable of
describing non-linear effects. A second major reason for the discrepancies
between the Equilibrium Climate Sensitivity and \ac{ECS} is the use of a
\nxcotwo{4} instead of a \nxcotwo{2} simulation. The factor of $2$ in the
denominator of \cref{eq:02:ecs} only partly compensates this difference since
the radiative forcing logarithmically depends on the atmospheric \ac{CO2}
concentration \autocite{Huang2014}.

@@

A further metric describing the climate sensitivity is the \ac{TCR}. In
contrast to the \ac{ECS}, this metric does not assume radiative equilibrium of
the Earth system but describes the transient response of an evolving climate.
Following \textcite{Bindoff2013}, \ac{TCR} is defined as the change in the
\ac{GSAT} at the time of \ac{CO2} doubling in a simulation with a \pct{1} per
year increase of the atmospheric \ac{CO2} concentration (\onepctcotwo{}). For
this, the annual mean \acp{GSAT} are averaged over a 20 year period centered at
the time of the \ac{CO2} doubling (years \range{61}{80} when the first year is
indexed with 1). To account for model drift, the annual mean changes in
\ac{GSAT} are calculated relative to a corresponding pre-industrial control
simulation smoothed with a linear fit that considers 140 model years (length of
the \onepctcotwo{} simulation). An illustration of that calculation is shown in
\cref{fig:02:tcr}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/tcr_calculation.pdf}
  \caption{Global and annual mean near-surface air temperature change $\Delta
    T$ for 140 years of a simulation with a \pct{1} per year increase of the
    atmospheric \ac{CO2} concentration (\onepctcotwo{}) for the \acf{CESM} of
    the \acf{NCAR} (blue circles). The temperature change is calculated
    relative to a corresponding pre-industrial control simulation smoothed
    with a linear fit over all 140 years. The \acf{TCR} is defined as $\Delta
    T$ at the time of \acs{CO2} doubling averaged over a 20 year period
    (illustrated by the vertical dashed lines). Here, $\text{TCR} = 2.35
    \unit{K}$ (horizontal dashed line).}
  \label{fig:02:tcr}
\end{figure}

Similar to \ac{ECS}, \ac{TCR} can also be defined in terms of an external
forcing and climate feedbacks. For this, the energy balance
\cref{eq:02:N_vs_delta_t} can be slightly adjusted. Since over \pct{90} of the
excess energy introduced into the system by the radiative forcing $F$ is taken
up by the ocean due to its large heat capacity, $N$ can be taken equal to the
global ocean heat uptake \autocite{Knutti2017}. In experiments with a steadily
increasing radiative forcing, which is the case for the \onepctcotwo{}
simulation, this ocean heat uptake can be approximated with
\begin{equation}
  N = \kappa \cdot \Delta T,
  \label{eq:02:ocean_heat_uptake}
\end{equation}
where $\kappa$ is the ocean heat uptake efficiency \autocite{Gregory2008}.
$\kappa > 0$, since there is a net energy flux into the climate system ($N >
0$) due to the external forcing $F$. This approximation becomes less accurate
as the deeper ocean warms up and cannot be applied to simulations with a steady
state climate change ($N \to 0$) where the external forcings tends stabilize
\autocite{Gregory2009}. By applying the definition of \ac{TCR} (transient
\ac{GSAT} change $\Delta T$ at the time of \ac{CO2} doubling) and combining
\cref{eq:02:N_vs_delta_t,eq:02:ocean_heat_uptake}, \ac{TCR} can be estimated as
\begin{equation}
  \text{TCR} = \frac{F_\text{2x}}{\kappa - \lambda},
  \label{eq:02:tcr}
\end{equation}
where $F_\text{2x}$ is the radiative forcing induced by a doubling of the
atmospheric \ac{CO2} concentration.

This equation can be used to derive a relationship between \ac{TCR} and
\ac{ECS}. Writing $\text{ECS} = -F_\text{2x} / \lambda$ and assuming path
independence of the forcing (\ie{} the resulting radiative forcing from the
\onepctcotwo{} and \nxcotwo{2} runs are comparable) gives
\begin{equation}
  \text{TCR} = \frac{F_\text{2x} \cdot \text{ECS}}{F_\text{2x} + \kappa \cdot
    \text{ECS}},
  \label{eq:02:tcr_vs_ecs}
\end{equation}
which demonstrates the non-linear connection between \ac{TCR} and \ac{ECS} \autocite{Gregory2008, Nijsse2020}.
